---
title: "Appendix: Power Analysis"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
library(lmerTest)
```

## Design


For this power analysis we will simulate 20 labs contributing 32 infants (640 participants) from 3 to 15 months of age.

Notes: MB1 overall effect size was 0.29 for the single-screen central fixation (CF) method, with additional effect of 0.21 for HPP, and eye-tracking (ET) yielding a slight (non-significant) decrease in effect of -0.06.
We expect to have X labs running infant-controlled familiarization duration, and the other 20-X labs running a fixed familiarization procedure.

Factors:

* *stimulus_type*: indicates the type (complexity/difficulty) of the stimuli that infants are familiarized with during training (high/low stimulus type; within-infant, 12 per type)

* *familiarization_time*: indicates how long each stimulus is exposed during familiarization (5, 10, or 15 seconds; within-infant) 

* *trial_num*: indicates the sequential order in which test trials were presented. Trial number thus ranges from 1 to 24. 

* *age_mos*: the infants' age in months (3.0-15.0), centered in *age* column.

* *procedure*: indicates the experimental method that was used to record infants’ looking to the stimuli: infant-controlled exposure (IC; total familiarization time is achieved over uncontrolled period of time) vs. fixed-duration exposure (FD; controlled period of exposure, unknown period of infant fixation)

* *test_order*: indicates which of the four pseudorandom test orders (from our provided scripts) were used to present test trials to the infant. true for MB5??? 

To do our power analysis, we will generate 1,000 datasets of this structure with a given effect size (e.g., .3), run the mixed-effects regression for each simulated dataset, and count the number of times that the effect is significant. 
Note that we generate normally-distributed looking times, assuming that they have already been log-transformed.

## Simulate Datasets

```{r simulate-data}
# To use the rbeta function, we need to supply two shape functions that correspond to the mean and the 
# variance of the data. To get those shape parameters, we will use the get.ab function below from 
# Mijke Rhemtulla/ Minhajuddin et al. (2004)
get.ab <- function(mu, var){
  v <- var
  w <- mu/(1-mu)
  b <- ((w/ (v*(w^2 + 2*w + 1))) - 1)  / (w + 1)
  a <- b*w
  return(list(a = a, b = b))
}

set.seed(123) # reproducible sampling

generate_dataset <- function(n_labs=20, n_per_lab=32, 
                             effect_sizes=list(type = .3, familiarization = .1, age = .1, "age*type"=.1,
                                               "age*familiarization"=0, "type*familiarization"=0,
                                               "type*age*familiarization"=.1)) { 
  # critical test is the 3-way interaction?
  
  # rewrite to use expand.grid ?
  labID = rep(LETTERS[1:n_labs], each=n_per_lab)
  subjID = 1:(n_labs*n_per_lab)

  familiarization_times = c(5,10,15)
  stimulus_types = c(rep("high",4), rep("low",4))
  # trials each subject gets (but randomly ordered)
  fam_by_stim = expand.grid(fam_time = familiarization_times, stimulus_type = stimulus_types)
  
  # assume each lab uses one procedure
  lab_procedure = sample(c("IC","FD"), n_labs, replace=T, prob=c(.5,.5)) # 50/50 IC / FD procedures?
  procedure = rep(lab_procedure, each=n_per_lab)

  test_order = rep(1:4, n_per_lab/4*n_labs) 
  
  # per-subject data
  simd <- tibble(subjID, labID, procedure, test_order) %>%
    mutate(subjInt = rnorm(length(subjID), mean=0, sd=1))

  # add lab random intercept
  simd$labInt = 0.0
  for(lab in unique(labID)) {
    labInd = which(simd$labID==lab)
    simd[labInd,]$labInt = rnorm(1, mean=0, sd=1) # could increase per-lab variability ..
  }
  
  # uniform random vars
  simd$age_mos = runif(nrow(simd), min=3.0, max=15.0)
  simd$age = scale(simd$age_mos, center=T, scale=T)[,1]
  
  # generate per-subject data, put in long (row per-trial) df
  
  siml <- tibble()
  for(i in 1:nrow(simd)) {
    # randomized trial order (but maybe should be done according to preset pseudorandom orders?)
    tmp_sdat <- fam_by_stim[sample(1:nrow(fam_by_stim), size=nrow(fam_by_stim), replace=F),]
    # need novel and familiar looking times, to calculate prop_novel -- 
    # UNLESS prop_novel turns out to be normally-distributed 
    # (which would be easier to generate, requiring fewer assumptions) - check empirical trial-level data
    stimulus_type = with(tmp_sdat, ifelse(stimulus_type=="high", .5, -.5)) 
    error_term = rnorm(nrow(tmp_sdat), 0, sd=1) + simd[i,]$labInt + simd[i,]$subjInt # add random slopes? (e.g. by age..)
    # rescale error to be >0
    # ToDo: scale familiarization time ?
    age_effect_subj = effect_sizes$age * rep(simd[i,]$age, nrow(tmp_sdat))
    
    # can we assume these are z-scored proportions of novel looking? maybe truncate them?
    # ToDo: check if problems when effect sizes are 0?
    tmp_sdat$dv_zscore = effect_sizes$type * stimulus_type + # main
      age_effect_subj +  # main
      effect_sizes$familiarization * tmp_sdat$fam_time +  # main
      effect_sizes$`age*type` * stimulus_type * effect_sizes$type * age_effect_subj + 
      effect_sizes$`age*familiarization` * age_effect_subj * tmp_sdat$fam_time * effect_sizes$familiarization + 
      effect_sizes$`type*familiarization` * tmp_sdat$fam_time * stimulus_type * effect_sizes$type + 
      effect_sizes$`type*age*familiarization` * stimulus_type * effect_sizes$type * age_effect_subj * tmp_sdat$fam_time * effect_sizes$familiarization + 
      error_term
    # standardize from normal means to beta mean: .5+/-.3 ??
    siml <- siml %>% 
      bind_rows(tmp_sdat %>% mutate(subjID = simd[i,]$subjID,
                                    labID = simd[i,]$labID,
                                    age = simd[i,]$age,
                                    age_mos = simd[i,]$age_mos,
                                    subjInt = simd[i,]$subjInt,
                                    labInt = simd[i,]$labInt,
                                    trial_num = 1:nrow(tmp_sdat)))
          #novel_looking_time = rnorm(n = nrow(tmp_sdat), mean=0, sd=1), # = .05
          #familiar_looking_time = rnorm(n = nrow(tmp_sdat), mean=0, sd=1), # = .05
          #prop_novel = novel_looking_time / (novel_looking_time + familiar_looking_time), # use beta distribution?
          #prop_novel = rbeta(n=nrow(tmp_sdat), shape1=??, shape2=??)
          # mean_beta = .5 + familiarization_time*age*type
               # how to choose beta parameters: more non-central = more of a novelty/familiarity effect
  }
  
  siml$trial_num_sc = scale(siml$trial_num, center=T, scale=T) 
  
  siml$subjID = as.factor(siml$subjID)
  # switch from dummy-code to effects code 
  siml$stimulus_type = as.factor(siml$stimulus_type)
  contrasts(siml$stimulus_type) = contr.sum(2)
  return(siml)
}

```

## Plot Example Dataset

We generate and plot an example dataset with stimulus *type* main effect size of .3, *age* main effect size of -.2, and an age*trial_type interaction effect size of .3.

```{r, fig.width=6, fig.height=4.5, caption="Log(looking time) by trial type and trial number with a simulated main effect (Cohen's d=.3) of trial type, age effect (d=-.2), and trial type * age interaction (d=.3). Shaded regions denote bootstrapped 95% confidence intervals."}

siml = generate_dataset(effect_sizes=list(type = .3, familiarization = .3, age = .3, "age*type"=.3,
                                               "age*familiarization"=.3, "type*familiarization"=.3,
                                               "type*age*familiarization"=.3))

d_sub <- siml %>% mutate(age_group = cut_interval(age_mos, length=3)) %>%
  group_by(subjID, stimulus_type, fam_time, age_group) %>%
  summarise(dv_zscore = mean(dv_zscore)) 

pos = position_dodge(width=.2)
# age group faceting
# dag <- siml %>% mutate(age_group = cut_interval(age_mos, length=3)) %>%
#   group_by(subjID, stimulus_type, fam_time, age_group) %>%
#   summarise(dv_zscore = mean(dv_zscore)) %>% 
#   group_by(stimulus_type, fam_time, age_group) %>%
#   tidyboot::tidyboot_mean(dv_zscore) # quite slow..
# 
# ggplot(dag, aes(x=fam_time, y=mean, group=stimulus_type, color=stimulus_type)) + 
#   facet_wrap(. ~ age_group) + 
#   ylab("Standardized proportion novel looking") + xlab("Familiarization Time") + 
#   geom_linerange(aes(ymin=ci_lower, ymax=ci_upper), pos=pos) + 
#   geom_point(data=d_sub, aes(x=jitter(fam_time), y=dv_zscore), alpha=.2) + 
#   theme_bw() + geom_smooth(method="lm")

dag <- siml %>% 
  group_by(subjID, stimulus_type, fam_time, age_mos) %>%
  summarise(dv_zscore = mean(dv_zscore)) %>% 
  group_by(stimulus_type, fam_time, age_mos) %>%
  tidyboot::tidyboot_mean(dv_zscore) 

ggplot(dag, aes(x=age_mos, y=mean, group=stimulus_type, color=stimulus_type)) + 
  facet_wrap(. ~ fam_time) + 
  geom_point(aes(y=mean, x=age_mos), pos=pos) + 
  ylab("Standardized proportion novel looking") + xlab("Age (months)") + 
  geom_linerange(aes(ymin=ci_lower, ymax=ci_upper), pos=pos) + 
  theme_bw() + geom_smooth(method="lm")
```


## Model Structure

Infants' proportion of looking at novel object (DV) ~ 1 + familiarization time (5, 10, 15) * stimulus type (high/low complexity) * age + (fam_time\*stim_type | subject) + (fam_time\*stim_type*age | lab) 
<!-- https://docs.google.com/document/d/1vwPOBSnVtrmOb-brNlQq3DcAEcXCG1p1gFfywhuwXK8/edit -->

```{r model}
# power for either just main effects, or just the 3-way 
main_efs <- c("stimulus_type1","age_mos","fam_time")
                  #"stimulus_type1:fam_time", "stimulus_type1:age_mos", "fam_time:age_mos",
                  #"stimulus_type1:fam_time:age_mos")

# power for just the 3-way
fit_simple_model <- function(siml) {
  m1 <- lmer(dv_zscore ~ 1 + stimulus_type * fam_time * age_mos + (1 | subjID) + (1 | labID), data=siml)
  return(summary(m1)$coefficients["stimulus_type1:fam_time:age_mos","Pr(>|t|)"]) # "Estimate","t value",
} 

fit_simple_model(siml)

# check both
fit_model <- function(siml) {
  m1 <- lmer(dv_zscore ~ 1 + stimulus_type * fam_time * age_mos + 
               (fam_time*stimulus_type | subjID) + (fam_time*stimulus_type | labID), data=siml)
  sig =c(#summary(m1)$coefficients["stimulus_type1","Pr(>|t|)"],
       summary(m1)$coefficients["stimulus_type1:fam_time:age_mos","Pr(>|t|)"])
  return(sig) # "Estimate","t value",
}

fit_model(siml)
# need to update fit_model to return significance of all desired effects (e.g., if effect_size$age!=0)
```

## Power Analysis

We use this simplified model for the power analysis:
y ~ 1 +  * stimulus_type * age * fam_time + (1 | subjID) + (1 | labID)

To do the power analysis, we simply generate 1000 datasets with main effect sizes of 0.1, 0.2, and 0.3 for trial type, age, and their interaction, run the above linear mixed-effects model, and report how many times 1) the trial type main effect and 2) the trial type * age interaction is significant.

```{r, power-analysis, message=F, warning=F}
# repeatedly generate data and  significance of trial_typesame
get_power <- function(effect_sizes, N=100, alpha=.05, verbose=F) {
  #p = data.frame(type=numeric(), "stimulus_type*age_mos*fam_time"=numeric())
  # parallelize
  p = rep(NA, N)
  #colnames(p) = c("stimulus_type","age_mos","fam_time","stimulus_type*age_mos*fam_time")
  for(i in 1:N) {
    p[i] = fit_simple_model(generate_dataset(effect_sizes=effect_sizes))
  }
  if(verbose) {
    #print(paste(length(which(p$type<alpha)), "of",N, "simulations had p <",alpha, "for trial type"))
    print(paste(length(which(p<alpha)), "of",N, "simulations had p <",alpha, "for stimulus_type*age_mos*fam_time"))
  }
  return(p)
}

N = 100

smol_effect_sizes=list(type = .1, familiarization = .1, age = .1, "age*type"=.1,
                                               "age*familiarization"=.1, "type*familiarization"=.1,
                                               "type*age*familiarization"=.1)
med_effect_sizes=list(type = .2, familiarization = .2, age = .2, "age*type"=.2,
                                               "age*familiarization"=.2, "type*familiarization"=.2,
                                               "type*age*familiarization"=.2)
large_effect_sizes=list(type = .3, familiarization = .3, age = .3, "age*type"=.3,
                                               "age*familiarization"=.3, "type*familiarization"=.3,
                                               "type*age*familiarization"=.3)

pvalues_pt1 = get_power(effect_sizes=smol_effect_sizes, N=N)

pvalues_pt2 = get_power(effect_sizes=med_effect_sizes, N=N)

pvalues_pt3 = get_power(effect_sizes=large_effect_sizes, N=N)
```

### Effect sizes = .1
`r paste(length(which(pvalues_pt1$type<.05)), "of",N, "simulations had p <",.05, "for trial type.")`
`r paste(length(which(pvalues_pt1[,"age*type"]<.05)), "of",N, "simulations had p <",.05, "for age*trial type.")` 
<!--(With 16 infants per lab (total N=320): 316 of 1000 simulations had p < 0.05 for trial type. 1000 of 1000 simulations had p < 0.05 for age*trial type.)-->

### Effect sizes = .2
`r paste(length(which(pvalues_pt2$type<.05)), "of",N, "simulations had p <",.05, "for trial type.")`
`r paste(length(which(pvalues_pt2[,"age*type"]<.05)), "of",N, "simulations had p <",.05, "for age*trial type.")` 
<!--(With 16 infants per lab (total N=320): 777 of 1000 simulations had p < 0.05 for trial type. 1000 of 1000 simulations had p < 0.05 for age*trial type.)-->

### Effect sizes = .3
`r paste(length(which(pvalues_pt3$type<.05)), "of",N, "simulations had p <",.05, "for trial type.")`
`r paste(length(which(pvalues_pt3[,"age*type"]<.05)), "of",N, "simulations had p <",.05, "for age*trial type.")` 
<!--(With 16 infants per lab (total N=320): 985 of 1000 simulations had p < 0.05 for trial type. 1000 of 1000 simulations had p < 0.05 for age*trial type.)-->

For context, .3 is the average effect size across all published developmental experiments.
(Any idea of the average empirical effect size (of age, complexity, or familiarization time) for habituation experiments??)