---
title: "Appendix: Power Analysis"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
library(lmerTest)
library(here)
```

## Design


For this power analysis we will simulate 40 labs contributing 32 infants (1280 participants) from 3 to 15 months of age. (MB5 estimates there will be a minimum of 1200 participants in the final sample.)

Notes: MB1 overall effect size was 0.29 for the single-screen central fixation (CF) method, with additional effect of 0.21 for HPP, and eye-tracking (ET) yielding a slight (non-significant) decrease in effect of -0.06.
We expect to have X labs running infant-controlled familiarization duration, and the other 20-X labs running a fixed familiarization procedure.

Factors:

* *stimulus_type*: indicates the type (complexity/difficulty) of the stimuli that infants are familiarized with during training (high/low stimulus type; within-infant, 12 per type)

* *familiarization_time*: indicates how long each stimulus is exposed during familiarization (5, 10, or 15 seconds; within-infant; scaled to [-.5,.5])

* *trial_num*: indicates the sequential order in which test trials were presented. Trial number thus ranges from 1 to 24. 

* *age_mos*: the infants' age in months (3.0-15.0), scaled and centered in *age* column.

* *procedure*: indicates the experimental method that was used to record infants’ looking to the stimuli: infant-controlled exposure (IC; total familiarization time is achieved over uncontrolled period of time) vs. fixed-duration exposure (FD; controlled period of exposure, unknown period of infant fixation)

<!--* *test_order*: indicates which of the four pseudorandom test orders (from our provided scripts) were used to present test trials to the infant. -->

To do our power analysis, we will generate 1,000 datasets of this structure with a given effect size (e.g., .3), run the mixed-effects regression for each simulated dataset, and count the number of times that the effect is significant. 
Note that we generate normally-distributed looking times, assuming that they have already been log-transformed.

## Simulate Datasets

```{r, include=F, echo=F}
# To use the rbeta function, we need to supply two shape functions that correspond to the mean and the 
# variance of the data. To get those shape parameters, we will use the get.ab function below from 
# Mijke Rhemtulla/ Minhajuddin et al. (2004)
get.ab <- function(mu, var){
  v <- var
  w <- mu/(1-mu)
  b <- ((w/ (v*(w^2 + 2*w + 1))) - 1)  / (w + 1)
  a <- b*w
  return(list(a = a, b = b))
}

```

```{r simulate-data}
set.seed(123) # reproducible sampling

generate_dataset <- function(n_labs=40, n_per_lab=32, 
                             effect_sizes=list(type = .3, 
                                               familiarization = .1,
                                               age = .1, "age*type"=.1,
                                               "age*familiarization"=0, "type*familiarization"=0,
                                               "type*age*familiarization"=.1)) { 
  # critical test is the 3-way interaction?
  
  # rewrite to use expand.grid ?
  labID = rep(as.character(1:n_labs), each=n_per_lab)
  subjID = 1:(n_labs*n_per_lab)

  familiarization_times = c(-0.5,0,0.5) # or maybe we expect linear effect on log(fam_time)?
  fam_times_sc = c(-0.5,0,0.5) # scaled
  #fam_times_sc = log(familiarization_times) # tried this: yields much lower power for main effects, only slight benefit for interactions
  stimulus_types = c(rep("high",4), rep("low",4)) # stimulus complexity
  # trials each subject gets (but randomly ordered)
  fam_by_stim = expand.grid(fam_time = fam_times_sc, stimulus_type = stimulus_types)
  
  # assume each lab uses one procedure
  lab_procedure = sample(c("IC","FD"), n_labs, replace=T, prob=c(.5,.5)) # 50/50 IC / FD procedures?
  procedure = rep(lab_procedure, each=n_per_lab)

  test_order = rep(1:4, n_per_lab/4*n_labs) 
  
  # per-subject data
  simd <- tibble(subjID, labID, procedure, test_order) %>%
    mutate(subjInt = rnorm(length(subjID), mean=0, sd=1))

  # add lab random intercept
  simd$labInt = 0.0
  for(lab in unique(labID)) {
    labInd = which(simd$labID==lab)
    simd[labInd,]$labInt = rnorm(1, mean=0, sd=1) # could increase per-lab variability ..
  }
  
  # uniform random vars
  simd$age_mos = runif(nrow(simd), min=3.0, max=15.0)
  simd$age = scale(simd$age_mos, center=T, scale=T)[,1]
  
  # generate per-subject data, put in long (row per-trial) df
  
  siml <- tibble()
  for(i in 1:nrow(simd)) {
    # randomized trial order (but maybe should be done according to preset pseudorandom orders?)
    tmp_sdat <- fam_by_stim[sample(1:nrow(fam_by_stim), size=nrow(fam_by_stim), replace=F),]
    # let's assume prop_novel is normally-distributed
    stimulus_type = with(tmp_sdat, ifelse(stimulus_type=="high", .5, -.5)) 
    error_term = rnorm(nrow(tmp_sdat), 0, sd=1) + simd[i,]$labInt + simd[i,]$subjInt # add random slopes? (e.g. by age..)
    # rescale error to be >0
    # ToDo: scale familiarization time ?
    age_effect_subj = effect_sizes$age * rep(simd[i,]$age, nrow(tmp_sdat))
    
    # can we assume these are z-scored proportions of novel looking? maybe truncate them?
    # ToDo: check if problems when effect sizes are 0?
    tmp_sdat$dv_zscore = effect_sizes$type * stimulus_type + # main
      age_effect_subj +  # main
      effect_sizes$familiarization * tmp_sdat$fam_time +  # main
      effect_sizes$`age*type` * stimulus_type * effect_sizes$type * age_effect_subj + 
      effect_sizes$`age*familiarization` * age_effect_subj * tmp_sdat$fam_time * effect_sizes$familiarization + 
      effect_sizes$`type*familiarization` * tmp_sdat$fam_time * stimulus_type * effect_sizes$type + 
      effect_sizes$`type*age*familiarization` * stimulus_type * effect_sizes$type * age_effect_subj * tmp_sdat$fam_time * effect_sizes$familiarization + 
      error_term
    
    # since DV has SD~.2, and must be in the range [0,1], let's make floor=-2.5 and ceiling=2.5. # the next four lines generate a     # strange distribution: 
    #min_ind = which(tmp_sdat$dv_zscore< -2.5)
    #max_ind = which(tmp_sdat$dv_zscore > 2.5)
    #if(length(min_ind)>0) tmp_sdat[min_ind,]$dv_zscore = -2.5
    #if(length(max_ind)>0) tmp_sdat[max_ind,]$dv_zscore = 2.5
    
    siml <- siml %>% 
      bind_rows(tmp_sdat %>% mutate(subjID = simd[i,]$subjID,
                                    labID = simd[i,]$labID,
                                    age = simd[i,]$age,
                                    age_mos = simd[i,]$age_mos,
                                    subjInt = simd[i,]$subjInt,
                                    labInt = simd[i,]$labInt,
                                    trial_num = 1:nrow(tmp_sdat)))
          #novel_looking_time = rnorm(n = nrow(tmp_sdat), mean=0, sd=1), # = .05
          #familiar_looking_time = rnorm(n = nrow(tmp_sdat), mean=0, sd=1), # = .05
          #prop_novel = novel_looking_time / (novel_looking_time + familiar_looking_time), # use beta distribution?
          #prop_novel = rbeta(n=nrow(tmp_sdat), shape1=??, shape2=??)
          # mean_beta = .5 + familiarization_time*age*type
               # how to choose beta parameters: more non-central = more of a novelty/familiarity effect
  }
  
  siml$trial_num_sc = scale(siml$trial_num, center=T, scale=T) 
  
  siml$subjID = as.factor(siml$subjID)
  # switch from dummy-code to effects code 
  siml$stimulus_type = as.factor(siml$stimulus_type)
  contrasts(siml$stimulus_type) = c(0.5, -0.5)
  
  return(siml)
}

```

## Plot Example Dataset

We generate and plot an example dataset with all effect sizes = .3 (main, 2-way, and 3-way).

```{r, fig.width=6, fig.height=4.5, echo=F, caption="z-scored proportion of novel looking by stimulus type by age, faceted by familiarization time. All simulated main and interaction effects are of size d=.3. Shaded regions denote bootstrapped 95% confidence intervals."}
effect_size_pt3 = list(type = .3, familiarization = .3, age = .3, "age*type"=.3,
                       "age*familiarization"=.3, "type*familiarization"=.3,
                       "type*age*familiarization"=.3)

siml = generate_dataset(effect_sizes = effect_size_pt3)

#plot things as a sanity check:
siml %>%
  ggplot(aes(stimulus_type, dv_zscore)) +
  geom_violin(aes(group = stimulus_type)) +
  geom_point(position = "jitter", size = 0.1, alpha = 0.2) +
  theme_bw()

  #save generated dataset
write.csv(siml, here::here("power_analysis","simulated_data","example_power_analysis_data_0.3.csv"), row.names=FALSE)
```


```{r, include=F, echo=F}
d_sub <- siml %>% mutate(age_group = cut_interval(age_mos, length=3)) %>%
  group_by(subjID, stimulus_type, fam_time, age_group) %>%
  summarise(dv_zscore = mean(dv_zscore)) 


dag <- siml %>% 
  group_by(subjID, stimulus_type, fam_time, age_mos) %>%
  summarise(dv_zscore = mean(dv_zscore)) %>% 
  group_by(stimulus_type, fam_time, age_mos) %>%
  tidyboot::tidyboot_mean(dv_zscore) %>%
  #mutate(fam_time = case_when(fam_time==log(5) ~ 5, # unscale
  #                            fam_time==log(10) ~ 10,
  #                            fam_time==log(15) ~ 15))
  mutate(fam_time = case_when(fam_time==-1 ~ 5, # unscale
                              fam_time==0 ~ 10,
                              fam_time==1 ~ 15))

ggplot(dag, aes(x=age_mos, y=mean, group=stimulus_type, color=stimulus_type)) + 
  facet_wrap(. ~ fam_time) + 
  geom_point(aes(y=mean, x=age_mos), alpha=.1) + 
  ylab("Standardized proportion novel looking") + xlab("Age (months)") + 
  geom_linerange(aes(ymin=ci_lower, ymax=ci_upper)) + 
  theme_bw() + geom_smooth(method="lm")
```

```{r, include=F, echo=F}
pos = position_dodge(width=.2)
#age group faceting
dag <- siml %>% mutate(age_group = cut_interval(age_mos, length=3)) %>%
  group_by(subjID, stimulus_type, fam_time, age_group) %>%
  summarise(dv_zscore = mean(dv_zscore)) %>%
  group_by(stimulus_type, fam_time, age_group) %>%
  tidyboot::tidyboot_mean(dv_zscore) # quite slow..

ggplot(dag, aes(x=fam_time, y=mean, group=stimulus_type, color=stimulus_type)) +
  facet_wrap(. ~ age_group) +
  ylab("Standardized proportion novel looking") + xlab("Familiarization Time") +
  geom_linerange(aes(ymin=ci_lower, ymax=ci_upper), pos=pos) +
  geom_point(data=d_sub, aes(x=jitter(fam_time), y=dv_zscore), alpha=.2) +
  theme_bw() + geom_smooth(method="lm")
```


## Model Structure

Infants' proportion of looking at novel object (DV) ~ 1 + familiarization time (-0.5, 0, .5) * stimulus type (high/low complexity) * age + (fam_time\*stim_type | subject) + (fam_time\*stim_type*age | lab) 
<!-- https://docs.google.com/document/d/1vwPOBSnVtrmOb-brNlQq3DcAEcXCG1p1gFfywhuwXK8/edit -->

```{r model}
# power for either just main effects, or just the 3-way 
effects <- c("stimulus_type1","age_mos","fam_time",
              "stimulus_type1:fam_time", "stimulus_type1:age_mos", "fam_time:age_mos",
              "stimulus_type1:fam_time:age_mos")

# power for just the 3-way
fit_simple_model <- function(siml) {
  m1 <- lmer(dv_zscore ~ 1 + stimulus_type * fam_time * age + (1 | subjID) + (1 | labID), data=siml)
  #coefs = as.data.frame(coef(summary(m1)))
  return(coef(summary(m1))[,5])
  #return(summary(m1)$coefficients[effects,"Pr(>|t|)"])
}

#fit_simple_model(siml)
fit_model <- function(siml) {
  m1 <- lmer(dv_zscore ~ 1 + stimulus_type * fam_time * age + (fam_time*stimulus_type | subjID) + (fam_time*stimulus_type | labID), data=siml)
  return(coef(summary(m1))[,5])
  #sig = summary(m1)$coefficients[effects,"Pr(>|t|)"]
  #return(sig)
}

#fit_model(siml) # boundary (singular) fit -- and is quite slow
```

## Power Analysis

We use this simplified model for the power analysis:
y ~ 1 +  * stimulus_type * age * fam_time + (1 | subjID) + (1 | labID)

To do the power analysis, we simply generate 1000 datasets with main effect sizes of 0.1, 0.2, and 0.3 for trial type, age, and their interaction, run the above linear mixed-effects model, and report how many times 1) the trial type main effect and 2) the trial type * age interaction is significant.

```{r, power-analysis, message=F, warning=F}
# repeatedly generate data and  significance of trial_typesame
get_power <- function(effect_sizes, N=100, alpha=.05) {
  p = tibble()
  # parallelize
  for(i in 1:N) {
    p <- p %>% bind_rows(fit_simple_model(generate_dataset(effect_sizes=effect_sizes)))
  }
  return(p)
}

N = 2

effect_size_pt1 = list(type = .1, familiarization = .1, age = .1, "age*type"=.1,
                       "age*familiarization"=.1, "type*familiarization"=.1,
                       "type*age*familiarization"=.1)
effect_size_pt2 = list(type = .2, familiarization = .2, age = .2, "age*type"=.2,
                       "age*familiarization"=.2, "type*familiarization"=.2,
                       "type*age*familiarization"=.2)

pvalues_pt1 = get_power(effect_sizes=effect_size_pt1, N=N)
pvalues_pt2 = get_power(effect_sizes=effect_size_pt2, N=N)
pvalues_pt3 = get_power(effect_sizes=effect_size_pt3, N=N)

report_main_effects <- function(pvalues) {
  paste(length(which(pvalues$stimulus_type1<.05)), "of",N, "simulations had p <",.05, "for stimulus type.",
        length(which(pvalues$age_mos<.05)), "of",N, "simulations had p <",.05, "for age.",
        length(which(pvalues$fam_time<.05)), "of",N, "simulations had p <",.05, "for familiarization time.")
}

report_interactions <- function(pvalues) {
  paste(length(which(pvalues[,"stimulus_type1:fam_time"]<.05)), "of",N, "simulations had p <",.05, "for stimulus type * familiarization time.",
        length(which(pvalues[,"stimulus_type1:age_mos"]<.05)), "of",N, "simulations had p <",.05, "stimulus type * age.",
        length(which(pvalues[,"fam_time:age_mos"]<.05)), "of",N, "simulations had p <",.05, "for familiarization time * age.",
        length(which(pvalues[,"stimulus_type1:fam_time:age_mos"]<.05)), "of",N, "simulations had p <",.05, "for age * stimulus type * familiarization time.")
}
```

### Effect sizes = .1
`r report_main_effects(pvalues_pt1)`
`r report_interactions(pvalues_pt1)`

<!-- 760 of 1000 simulations had p < 0.05 for stimulus type.  850 of 1000 simulations had p < 0.05 for age. 987 of 1000 simulations had p < 0.05 for familiarization time. 47 of 1000 simulations had p < 0.05 for age * stimulus type * familiarization time. -->


### Effect sizes = .2
`r report_main_effects(pvalues_pt2)`
`r report_interactions(pvalues_pt2)`
<!-- 998 of 1000 simulations had p < 0.05 for stimulus type. 1000 of 1000 simulations had p < 0.05 for age. 1000 of 1000 simulations had p < 0.05 for familiarization time. 45 of 1000 simulations had p < 0.05 for age *stimulus type * familiarization time. -->


### Effect sizes = .3
`r report_main_effects(pvalues_pt3)`
`r report_interactions(pvalues_pt3)`
<!-- 1000 of 1000 simulations had p < 0.05 for stimulus type. 1000 of 1000 simulations had p < 0.05 for age. 1000 of 1000 simulations had p < 0.05 for familiarization time. 81 of 1000 simulations had p < 0.05 for age *stimulus type * familiarization time. -->

For context, .3 is the average effect size across all published developmental experiments.
(Any idea of the average empirical effect size (of age, complexity, or familiarization time) for habituation experiments??)