---
title: "Appendix: Power Analysis"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
library(lmerTest)
```

## Design


For this power analysis we will simulate 20 labs contributing 20 infants (400 participants) from 5 to 12 months of age.

Notes: MB1 overall effect size was 0.29 for the single-screen central fixation (CF) method, with additional effect of 0.21 for HPP, and eye-tracking (ET) yielding a slight (non-significant) decrease in effect of -0.06.
We expect to have 5-6 labs running HPP (100-120 infants), and the other 15-20 labs running CF/ET.

Factors:

* *familiarization time/ time on stimulus*: 5, 10, 15 s

* *total time*: total trial time - can't be less

* *stimulus complexity*: high, low

* *trial_num*: trial num ranges from 1-24

* *age_mos*: the infants' age in months (3.0-15.0), centered in *age* column.

* *procedure*: indicates the experimental method that was used to record infants’ responses to the stimuli: infant-controlled vs. fixed presentation time

* *test_order*: 
 

To do our power analysis, we will generate 1,000 datasets of this structure with a given effect size (e.g., .3), run the mixed-effects regression for each simulated dataset, and count the number of times that the effect is significant. 
Note that we generate normally-distributed looking times, assuming that they have already been log-transformed.

## Simulate Datasets

```{r simulate-data}
set.seed(123) # reproducible sampling

generate_dataset <- function(n_labs=20, n_per_lab=20, effect_sizes=list(type = .3, age = 0, "age*type"=0)) {
  # rewrite to use expand.grid ?
  labID = rep(LETTERS[1:n_labs], each=n_per_lab)
  subjID = 1:(n_labs*n_per_lab)

  # assume each lab uses one procedure
  lab_procedure = sample(c("HPP","CF","EF"), n_labs, replace=T, prob=c(.5,.3,.2))
  procedure = rep(lab_procedure, each=n_per_lab)

  test_order = rep(1:4, 5*n_labs) 

  # familiarized rule (ms says randomly assigned: we don't want counterbalanced per lab?)
  familiarized_rule = sample(c("ABB","ABA"), length(subjID), replace=T)

  simd <- tibble(subjID, labID, procedure, test_order, familiarized_rule)

  # uniform random vars
  simd$age_mos = runif(nrow(simd), min=5.0, max=12.0)
  simd$age = scale(simd$age_mos, center=T, scale=T)[,1]

  # should actually be bimodal (use MB1 distro?)
  simd$multilingual_exposure = runif(nrow(simd), min=0, max=.5) # 0=monolingual, .5=50% secondary language

  # now generate looking times for 12 trials per subject
  for(t in 1:12) {
    simd[,paste0("trial.",t)] = rnorm(n = nrow(simd), mean=0, sd=1) # = .05
  } 
  
  siml <- simd %>% pivot_longer(cols=starts_with("trial."), 
                     names_to="trial_num", 
                     names_prefix="trial.",
                     values_to="looking_time") 

  siml$trial_num = as.numeric(siml$trial_num)
  siml$trial_num_sc = scale(siml$trial_num, center=T, scale=T) 

  # 6 same / 6 different per child; should be according to 1 of 4 pseudorandom orders, but we're not  actually modeling order effects here so just make blocks:
  siml$trial_type = rep_len(c(rep("same", 6), rep("different", 6)), nrow(siml)) # each

  per_subj_trial_type = c(rep("same", 6), rep("different", 6))
  # add subject random intercept
  siml$subjInt = 0.0
  for(s in 1:length(unique(siml$subjID))) {
    subjInd = which(siml$subjID==s)
    siml[subjInd,]$trial_type = sample(per_subj_trial_type, 12, replace = F)
    siml[subjInd,]$subjInt = rnorm(1, mean=0, sd=1)
  }
  
  # add lab random intercept
  siml$labInt = 0.0
  for(lab in labID) {
    labInd = which(siml$labID==lab)
    siml[labInd,]$labInt = rnorm(1, mean=0, sd=1) # could increase per-lab variability ..
  }
  
  trial_type = with(siml, ifelse(trial_type=="same", 0, 1)) 
  error_term = rnorm(nrow(siml), 0, sd=1) + siml$labInt + siml$subjInt 
  siml$looking_time = trial_type * effect_sizes$type + siml$age * effect_sizes$age + trial_type * siml$age * effect_sizes$`age*type` + error_term
  
  siml$subjID = as.factor(siml$subjID)
  # switch from dummy-code to effects code 
  siml$familiarized_rule = as.factor(siml$familiarized_rule)
  siml$trial_type = as.factor(siml$trial_type)
  contrasts(siml$familiarized_rule) = contr.sum(2)
  contrasts(siml$trial_type) = contr.sum(2)
  return(siml)
}

```

## Plot Example Dataset

We generate and plot an example dataset with trial_type main effect size of .3, age main effect size of -.2, and an age*trial_type interaction effect size of .3.

```{r, fig.width=6, fig.height=4.5, caption="Log(looking time) by trial type and trial number with a simulated main effect (Cohen's d=.3) of trial type, age effect (d=-.2), and trial type * age interaction (d=.3). Shaded regions denote bootstrapped 95% confidence intervals."}
#siml = generate_dataset(effect_sizes=list(type = .3, age = 0, "age*type"=0))
siml = generate_dataset(effect_sizes=list(type = .3, age = -.2, "age*type"=.3))

dag <- siml %>% group_by(subjID, trial_type, age_mos) %>%
  summarise(looking_time = mean(looking_time)) %>% 
  group_by(trial_type, age_mos) %>%
  tidyboot::tidyboot_mean(looking_time) # quite slow..

pos = position_dodge(width=.2)
ggplot(dag, aes(x=age_mos, y=mean, group=trial_type, color=trial_type)) + 
  geom_point(aes(y=mean, x=age_mos), pos=pos) + 
  ylab("Standardized log(looking time)") + xlab("Age (months)") + 
  geom_linerange(aes(ymin=ci_lower, ymax=ci_upper), pos=pos) + 
  theme_bw() + geom_smooth(method="lm")
```


## Model Structure

Infants' log(looking time) (DV) ~ 1 + familiarization order (ABB vs ABA) * trial_type + age * trial_type (same rule vs different rule at test) + experimental_method (HPP vs central fixation vs eye-tracking) * trial_type + multilingual_exposure * trial_type + trial_num * trial_type + (trial_num*trial_type | subject) + (test_order | lab)

```{r model}
# m1 <- lmer(looking_time ~ 1 + trial_type * 
#              (familiarized_rule + age + procedure + multilingual_exposure + trial_num) +
#              (trial_num * trial_type | subjID) + (test_order | labID), data=siml)

# model without age
fit_simple_model <- function(siml) {
  m1 <- lmer(looking_time ~ 1 + trial_type * trial_num_sc + (1 | subjID), data=siml)
  return(summary(m1)$coefficients["trial_type1","Pr(>|t|)"]) # "Estimate","t value",
} # trial_type1 = different

# check both
fit_model <- function(siml) {
  m1 <- lmer(looking_time ~ 1 + trial_type * trial_num_sc + trial_type * age + (1 | subjID) + (1 | labID), data=siml)
  sig =c(summary(m1)$coefficients["trial_type1","Pr(>|t|)"],
       summary(m1)$coefficients["trial_type1:age","Pr(>|t|)"])
  return(sig) # "Estimate","t value",
}

# need to update fit_model to return significance of all desired effects (e.g., if effect_size$age!=0)
```

## Power Analysis

We use this simplified model for the power analysis:
y ~ 1 + trial_type * trial_num + trial_type * age + (1 | subjID) + (1 | labID)

To do the power analysis, we simply generate 1000 datasets with main effect sizes of 0.1, 0.2, and 0.3 for trial type, age, and their interaction, run the above linear mixed-effects model, and report how many times 1) the trial type main effect and 2) the trial type * age interaction is significant.

```{r, power-analysis, message=F, warning=F}
# repeatedly generate data and  significance of trial_typesame
get_power <- function(effect_sizes, N=100, alpha=.05, verbose=F) {
  p = data.frame(type=numeric(), "age*type"=numeric())
  colnames(p) = c("type","age*type")
  for(i in 1:N) {
    p[i,] = fit_model(generate_dataset(effect_sizes=effect_sizes))
  }
  if(verbose) {
    print(paste(length(which(p$type<alpha)), "of",N, "simulations had p <",alpha, "for trial type"))
    print(paste(length(which(p[,"age*type"]<alpha)), "of",N, "simulations had p <",alpha, "for age*trial type"))
  }
  return(p)
}

N = 1000
pvalues_pt1 = get_power(effect_sizes=list(type = .1, age = .1, "age*type"=.1), N=N)

pvalues_pt2 = get_power(effect_sizes=list(type = .2, age = .2, "age*type"=.2), N=N)

pvalues_pt3 = get_power(effect_sizes=list(type = .3, age = .3, "age*type"=.3), N=N)
```

### Effect sizes = .1
`r paste(length(which(pvalues_pt1$type<.05)), "of",N, "simulations had p <",.05, "for trial type.")`
`r paste(length(which(pvalues_pt1[,"age*type"]<.05)), "of",N, "simulations had p <",.05, "for age*trial type.")` 
<!--(With 16 infants per lab (total N=320): 316 of 1000 simulations had p < 0.05 for trial type. 1000 of 1000 simulations had p < 0.05 for age*trial type.)-->

### Effect sizes = .2
`r paste(length(which(pvalues_pt2$type<.05)), "of",N, "simulations had p <",.05, "for trial type.")`
`r paste(length(which(pvalues_pt2[,"age*type"]<.05)), "of",N, "simulations had p <",.05, "for age*trial type.")` 
<!--(With 16 infants per lab (total N=320): 777 of 1000 simulations had p < 0.05 for trial type. 1000 of 1000 simulations had p < 0.05 for age*trial type.)-->

### Effect sizes = .3
`r paste(length(which(pvalues_pt3$type<.05)), "of",N, "simulations had p <",.05, "for trial type.")`
`r paste(length(which(pvalues_pt3[,"age*type"]<.05)), "of",N, "simulations had p <",.05, "for age*trial type.")` 
<!--(With 16 infants per lab (total N=320): 985 of 1000 simulations had p < 0.05 for trial type. 1000 of 1000 simulations had p < 0.05 for age*trial type.)-->

For context, .25 is the average effect size from the meta-analysis of rule learning, and .3 is the average effect size across all published developmental experiments.
Thus, the latter two power simulations probably pertain in our case.