---
title: "Peekbank RT Demo"
author: "Martin Zettersten & the Peekbank Team"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

# Download peekbankr

If you haven't downloaded peekbankr (https://github.com/langcog/peekbankr) yet, be sure to do so first by uncommenting the lines below.

```{r}
# install.packages("remotes") # can also use devtools
# remotes::install_github("langcog/peekbankr")
```


# Preliminaries

Load packages. 

```{r}
LOAD_ALL_DATA <- TRUE

library(peekbankr)
library(tidyverse)
library(here)

knitr::opts_chunk$set(cache = TRUE, arn = FALSE,warning=FALSE, message = FALSE,cache.lazy = FALSE)
```

# Data Loading

1. Connect to the peekbank database
2. Pull individual tables using peekbankr functions ("get_[name_of_table]"). For details on the specific tables, including a codebook, see the peekbank documentation: https://peekbank.stanford.edu/docs/documentation/
3. Join these tables together to get a single tidy dataset in "long" format (each row is an observation at a particular timepoint)

```{r, eval = LOAD_ALL_DATA}
#connect to the database
con <- connect_to_peekbank()
#get all of the tables you need
datasets <- get_datasets(connection = con) %>% collect()
administrations <- get_administrations(connection = con) %>% collect()
subjects <- get_subjects(connection = con) %>% collect()
stimuli <- get_stimuli(connection = con) %>% collect()
trial_types <- get_trial_types(connection = con) %>% collect()
trials <- get_trials(connection = con)  %>% collect()
aoi_timepoints <- get_aoi_timepoints(connection = con) %>% collect()

aoi_data_joined <- aoi_timepoints %>%
  right_join(administrations) %>%
  right_join(subjects) %>%
  right_join(trials) %>%
  right_join(trial_types) %>%
  right_join(datasets) %>%
  mutate(target_stimulus_id = target_id) %>%
  right_join(rename_with(stimuli, ~ paste("target",.x , sep = "_"))) %>%
  mutate(distractor_stimulus_id = distractor_id) %>%
  right_join(rename_with(stimuli, ~ paste("distractor",.x , sep = "_")))

saveRDS(aoi_data_joined,here("power_analysis","peekbank","data","aoi_data_joined.Rds"), )
```


